---
title: "Simultaneous Dimensionality and Complexity Model Selection for Spectral Graph Clustering"
author: "Department of Applied Mathematics and Statistics, Johns Hopkins University"
date: '`r Sys.Date()`'
output:
  html_document:
    highlight: pygments
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

h1.title {
  color: DarkRed;
}
h1 { /* Header 1 */
  color: DarkBlue;
}
h2 { /* Header 2 */
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
div.figure {
  color: black;
  font-weight: bold;
  font-size: 1.0em;
} 
</style>


```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.align = 'center', message = FALSE)
```

This is a simple demo written in `R` to compare performances of graph clustering algorithms used in the [paper](https://arxiv.org/abs/1904.02926):

> Congyuan Yang, Carey E. Priebe, Youngser Park, David J. Marchette,  
Simultaneous Dimensionality and Complexity Model Selection for Spectral Graph Clustering,  
_Journal of Computational and Graphical Statistics_, in revision, 2020.

To run a simple working live demo, please run these lines

```{r run, eval=FALSE}
if (!require(RCurl)) install.packages("RCurl")
library(RCurl)

script <- getURL("https://raw.githubusercontent.com/youngser/dhatkhat/master/R/demo.R", ssl.verifypeer = FALSE)
eval(parse(text = script))
```

and the result (each plot on its own window) should look like below.   
__(This may take around 10 minutes on a typical laptop.)__

# Install/Load required packages and utility functions

```{r pkg}
## required packages and sources
options(repos=structure(c(CRAN="http://cloud.r-project.org/")))
requiredPkg <- c("tidyverse", "igraph", "Matrix", "mclust", "irlba")
newPkg <- requiredPkg[!(requiredPkg %in% installed.packages()[, "Package"])]
if (length(newPkg)) install.packages(newPkg, dependencies = TRUE)
update.packages(newPkg, ask = FALSE)
invisible(sapply(requiredPkg, require, character.only = TRUE))

util.file <- url("http://www.cis.jhu.edu/~parky/dhatKhat/R/utils.R")
source(util.file)
close(util.file)
```

# Set parameters

This is exactly the same setting as Figures 8 and 10 in the paper (with $p = 0.09$) except that the graph is smaller, e.g., $n = 100$ as opposed to $n = 500$ in the paper.

```{r param}
## parameters
n <- 100 # number of vertices
p <- 0.09 # a probability that between blocks can have edges
B <- cbind( c(.2, p), c(p, .1) ) # blockconnectivityprobabilitymatrix
rho <- c(.5, .5) # block membership probabilities, \pi in the paper
K <- length(rho) # number of blocks
Y <- rep(1:K, times=rho*n) # block membership vector, vertex labels
```

# Generate graph

```{r g}
## generate a graph
set.seed(42+49182)
g <- sbm.game(n, pref.matrix = as.matrix(B),
             block.sizes = rho*n, directed = F, loops = F); summary(g)
#image(as_adjacency_matrix(g), main="Adjacency Matrix")
#dev.new(); 
plotA(g)
```

# Embed graph

```{r emb}
## embed graph
dmax <- 8 # maximum embedding dimension
ase <- embed.graph(A=g[,,sparse=F], dim=dmax, scaling = TRUE)
#dev.new()
elb <- getElbows(abs(ase$D), n=3, plot=TRUE); title("Scree plot with 3 suggested elbows")
```

# Cluster graph

(Note that it may take ~30 seconds to run on a typical personal computer.)

```{r mc}
## run our algorithms
Kmax <- 4 # maximum number of clusters
bic <- ari <- rep(-Inf, dmax-1)
Khat <- rep(0, dmax-1)
out.SMS <- NULL; out.SMS$bic <- -Inf
for (d in 1 : (dmax - 1)) {
  for (k in 1 : Kmax) {
    # SMS_Reduced
    system.time(mout.SMS_R <- bicGmmVVV(ase$X[, 1:d, drop=FALSE], k))
    if (mout.SMS_R$bic > bic[d]) {
      bic[d] <- mout.SMS_R$bic
      ari[d] <- adjustedRandIndex(Y, mout.SMS_R$classification)
      Khat[d] <- k
    }

    # SMS
    system.time(mout.SMS <- bicGmmBlock2(ase$X, d, k))
    if (mout.SMS$bic > out.SMS$bic) {
      out.SMS <- mout.SMS
    }
  }
}

## calculate ARIs
## SMS
ari_SMS <- adjustedRandIndex(Y, out.SMS$classification)
dhat_SMS <- out.SMS$d
khat_SMS <- out.SMS$k

## SMS_Reduced
ari_SMS_R <- ari[dhat_SMS]
dhat_SMS_R <- dhat_SMS
khat_SMS_R <- khat_SMS

## ZG1
dhat_ZG1 <- min(elb[1], dmax-1)
ari_ZG1 <- ari[dhat_ZG1]
khat_ZG1 <- Khat[dhat_ZG1]

## ZG2
dhat_ZG2 <- min(elb[2], dmax-1)
ari_ZG2 <- ari[dhat_ZG2]
khat_ZG2 <- Khat[dhat_ZG2]

## ZG3
dhat_ZG3 <- min(elb[3], dmax-1)
ari_ZG3 <- ari[dhat_ZG3]
khat_ZG3 <- Khat[dhat_ZG3]
```

# Try other algorithms

(Note that it may take ~600 seconds to run on a typical personal computer.)

```{r alg}
## Louvain
result_Louvain <- cluster_louvain(g)
ari_Louvain <- adjustedRandIndex(Y, result_Louvain$membership)
khat_Louvain <- max(membership(result_Louvain))

## Walktrap
result_Rw <- cluster_walktrap(g)
ari_Walktrap <- adjustedRandIndex(Y, result_Rw$membership)
khat_Walktrap <- max(membership(result_Rw))

## IRM
if (!file.exists("RData/irm_out.RData")) {
  system.time(result_IRM <- irm(g[,,sparse=FALSE], sweeps = 1000)) # ~10 min to run on a macbook
  ari_IRM <- max(sapply(1:nrow(result_IRM), function(x) adjustedRandIndex(Y, result_IRM[x,])))
  khat_IRM <- max(mode.irm(result_IRM))
  dir.create("RData", showWarnings = FALSE)
  save(result_IRM, ari_IRM, khat_IRM, file="RData/irm_out.RData")
} else {
  load("RData/irm_out.RData")
}
```

# Print/Plot Output

```{r output}
algs <- c("SMS", "SMS_R", "ZG1", "ZG2", "ZG3", "Louvain", "Walktrap", "IRM")
df.out <- data.frame(method = factor(algs, levels=algs),
                     ARI = c(ari_SMS, ari_SMS_R, ari_ZG1, ari_ZG2, ari_ZG3,
                             ari_Louvain, ari_Walktrap, ari_IRM),
                     dhat = c(dhat_SMS, dhat_SMS_R, dhat_ZG1, dhat_ZG2, dhat_ZG3,
                              NA, NA, NA),
                     Khat = c(khat_SMS, khat_SMS_R, khat_ZG1, khat_ZG2, khat_ZG3,
                              khat_Louvain, khat_Walktrap, khat_IRM))
df.out

p <- df.out %>% ggplot(aes(method, ARI, color=method, fill=method)) +
  geom_col(alpha=0.5, show.legend = FALSE) +
  geom_text(aes(label=paste0("hat(K) == ", Khat)), parse=TRUE, vjust=-0.2, show.legend = FALSE) +
  theme(axis.title=element_text(size=15)) +
  theme(axis.text.x=element_text(size=12)) +
  #  theme(axis.ticks = element_line(size = 5)) +
  theme(axis.text.y=element_text(size=15)) +
  theme(strip.text=element_text(size=rel(1.2))) +
  theme(legend.text = element_text(colour="black", size = 15, face = "plain"))
#grid::grid.newpage()
print(p)
```
